### 2020-03-04 Common/chapters/deep_learning.tex
Johan and Julien: all names for each section added

\chapter{Deep learning}
\label{ch:deep_learning}
This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work. It starts with a historical timeline of deep learning before describing what a neural network is. Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained. This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks". They are used in many computer vision applications due to their great performance on these tasks. Finally, the concept of "transfer learning" is discussed as the entire chapter \ref{ch:transfer_learning} relies on it.

\section{Introduction to deep learning}
Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence. Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them. In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way. There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc. What mainly differentiates deep learning from these algorithms is the concept of neural networks (see Section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics etc. The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology. Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

\subsection{Historical background}
\label{historical_background}
As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers. From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks. One could ask why deep learning took off recently if the theory was around for a long time.\\
As stated by Goodfellow et al. \cite{15}, the first part of the answer is computing power. In fact, deep learning algorithms need a lot of data to work properly, which requires powerful CPUs/GPUs that either didn't exist or were only within few people's reach. One other main reason concerns the lack of data. Since deep learning algorithms "learn" from data, learning is impossible if good-quality data is not available. The era of Big Data enhanced deep learning possibilities. 
%%%% Je mettrais soit l'un soit l'autre. Le paragraphe du dessus c'est le paragraphe du dessous avec des autres mots :)
% These two points are summarised: "The increase in model size overtime, due to the availability of faster CPUs, the advent of general purpose GPUs, fasternetwork, connectivity and better software infrastructure for distributed computing, is one of the most important trends in the history of deep learning. This trend is generally expected to continue well into the future".\\
Finally, before the year 2012, the abilities of neural networks were still to be proven. This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluate their algorithms on several visual recognition tasks). In fact, the deep convolutional neural network called "AlexNet" achieved 16\% of classification error rate, whereas the previous best scores were around 25\%. This victory marked the beginning of a new craze for these types of algorithms.

@@

\subsection{What is a neural network?}
\label{what_is_a_neural_network}

From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters. The input can be an image, a sound or any input that can be converted into numerical features. The output of a neural network depends on the problem it tries to solve. In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
\begin{equation}
y = f(x,\theta)
\end{equation}
The parameters $\theta$ are estimated from the training samples.
%\begin{figure}[!h]
%\centering
%\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/neural_network.png}
%\caption{Neural Network example}
%\label{neural_network}
%\end{figure}

\subsection{Supervised and unsupervised learning}
Machine learning algorithms belongs can be classified in two classes. The first one is "supervised learning". It includes learning algorithms whose training samples are associated with their labels in order to find the optimal mapping between the input and the output. The second one is "unsupervised learning". In contrast to supervised algorithms, the latter rely on unlabeled data. Its main goal is to infer the natural structure  present in the data. Since the models presented in this work belong to the "supervised learning" category, notions explained below refer to this kind of algorithms.

@@

\subsection{Notation}
In order to keep the mathematical description of neural network consistent, this work will use Andrew Y. Ng's notation \cite{16}, who was a pioneer in deep learning.\\

@@

\subsection{Perceptrons}
\label{perceptron}
Perceptrons are the main components of neural networks. They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}. Today, they are called "artificial neurons" or simply "neurons".\\
The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
\begin{equation}
\label{perceptron_equation}
a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{j})
\end{equation}

@@

\subsection{Activation functions}
\label{activation_functions}
Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer. Activation functions must be non-linear in order to approximate extremely complex functions. In fact, neural networks are considered as universal approximators. Hornik et al. claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}. According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
 \noindent \textbf{Sigmoid function}\\
 The sigmoid function is defined as:
 \begin{equation}
  g(x) = \frac{1}{1+e^{-x}}
 \end{equation}
 Its derivative is:
 \begin{equation}
 g'(x) = g(x)(1-g(x))
 \end{equation}

@@

\subsection{Multilayer perceptrons}
\label{multilayer_perceptron}
A multilayer perceptron is a type of artifical neural networks. Du et al. \cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
input and output layers" where "the output units represent a hyperplane in the space
of the input patterns."\\
A multilayer perceptron is composed of $L$ layers, each of them composed of $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
\begin{itemize}
\item The input layer: It is the neural network entry point for the data. Generally speaking, the data is provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$. The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is the dimension of each sample. Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector. The total number of training samples is given by $m$. Finally, $y^{(i)}$ is the output label for the $i^{th}$ example. For instance, suppose the number of samples is 100 and the batch size 32. In this situation, the network will be fed with 4 batches of sizes [32, 32, 32, 4] respectively.

@@

\section{Training a neural network}
\label{training_a_neural_network}
Training a neural network can be broken down into multiple steps. The first one is the "forward propagation" step. It consists in giving examples that need to be classified (or segmented, depending on the task) to the untrained neural network and to spread intermediate results through all layers of the network. %Passing all batches as input to a network is called an "epoch". 
After seeing every single batch, the loss is computed using a "loss function". The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth. Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minimum of the loss function. As soon as the neural network has seen every single batch, the end of an "epoch" is reached. This process is repeated for a defined number of epochs.

\subsection{Forward propagation}
The forward propagation is used to transmit the input through the entire neural network. Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations. The first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$: 
\begin{equation}
z^{[l]} = W_{x}^{[l]}x^{(i)} + b^{[l]}
\end{equation}
The second equation describes the effect of the activation function:
\begin{equation}
a^{[l]} = g^{[l]}(z^{[l]})
\end{equation}
Since the output of the activation function is then given as input to all the neurons of the next layer, the whole forward propagation step can be defined as:
\begin{equation}
a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) 
\end{equation}
Figure \ref{forward_propagation} illustrates the computation of the forward propagation pass for the $l^{th}$ layer. The weight natrix $W_{jk}^{l}$ represents the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$. This matrix is multiplied by the output of the previous layer $a_{j}^{[l-1]}$ before adding the bias $b^{[l]}$. The result is given as input to the activation function.

@@

\subsection{Loss computation}
As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}. It is therefore useful to measure the penalty for a single input. On the contrary, when the goal is to get a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used. The latter is represented by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.

@@

\subsection{Backpropagation}
Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$. Generally speaking, "the intuition behind the backpropagation algorithm is as follows. Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network. Then, for each node $i$ in layer $l$, we would like to compute an "error term" $\partial^{(l)}_{i}$ that measures how much that node was "responsible" for any errors in our output. For an output node, we can directly measure the difference between the network’s activation and the true target value, and use that to define $\partial^{(n_{l})}_{i}$(where layer $n_{l}$ is the output layer). How about hidden units? For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.

@@

\subsection{Metrics}
In classification tasks, four separate output labels can occur:
\begin{itemize}
\item True Positive (TP):  an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{correct}.
\item True Negative (TN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{correct}.
\item False Positive (FP): an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{incorrect}.
\item False Negative (FN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{incorrect}.
\end{itemize}
From these four categories, multiple metrics with their own specificities can be computed \cite{25}:
\begin{itemize}
\item Accuracy: "Ratio of the correctly labeled subjects to the whole pool of subjects".
\begin{equation}
Accuracy = \frac{(TP+TN)}{TP+FP+FN+TN}
\end{equation}
Accuracy is a great measure in the case of symmetric data (i.e the number of FN $\approx$ FP and their cost is similar). When this condition is not fulfilled, accuracy can lead to bad models. For instance, let's define a binary classification model that always outputs class 0. If the data is composed of 99 samples from class 0 and 1 sample from class 1, the accuracy is equal to 99\% but the model is not smart. Consequently, this metric has to be used in addition to other metrics.

@@

\subsection{Data}
In deep learning, data is essential. As seen previously, neural networks learn features from it. Therefore, data has to be handled carefully and in the right way. Usually, it is split into three different sets:
\begin{itemize}
\item The training set: This is the part of the dataset that is used to train the neural network (the weights and biases).
\item The validation set: This dataset is used to evaluate a trained model. Usually, the evaluation on the validation set is performed every $N$ epochs, where $N$ is a fixed number. The validation set needs to come from the same distribution as the training set but should contain exclusively unseen data. This last point is crucial since the validation set shows how well the neural network generalizes on unknown data. The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" which is the behaviour of a model that fits to the training set too closely and don't generalize well. In fact, if the validation loss continuously increases for a certain number of epochs, going on with the training will increase the overfit. This fact of interrupting the training earlier is called "early stopping".
\item The test set: This last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
\end{itemize}
Regarding the way these three sets are split, it mostly depends on the number of samples available. If the latter is big, the data is split into training and testing using the ratio 80/20. Then the remaining training samples are also split into training and validation using the ratio 80/20. On the contrary, if there are few data available, k-fold cross-validation is a good practice. This technique consists in splitting the entire dataset into k folds. One fold is picked as test set and the others are considered as training sets. The model is trained on training folds and tested on the test set. Then, another test set is picked and the same process is repeated until all possible tests set are picked. At the end of the process, the results of all test sets are averaged, which provides a good estimation of the model's performance. This technique is summarized on figure \ref{cross_validation}.

@@

\subsection{Weight initialization}
\label{weight_initialization}
Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation. The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum. The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network. If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.\\
The simplest and least efficient technique to initialize neural network weights is to randomly generate them. The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions. Consequently, the slope of the gradient changes slowly and the learning takes a lot of time.\\
To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
\begin{equation}
\frac{\sqrt{6}}{\sqrt{n_{h}^{[l]}+n_{h}^{[l+1]}}}
\end{equation}
where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
For activation functions that are not symmetric around zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative. It consists in multiplying the randomly initialized  weight matrix by:
\begin{equation}
\frac{\sqrt{2}}{\sqrt{n_{h}^{[l]}}}
\end{equation} 
where $n_{h}^{[l]}$ is the number of incoming connections coming into a given layer from the previous layer's output.

\subsection{Hyperparameters tuning}
Hyperparameters denote parameters that cannot be directly learned from the data. That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections. So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).\\
Regarding the learning rate $\alpha$, its value has to be neither too large nor too small. A too large learning rate is recognizable by analyzing the training loss curve: if the loss is exploding or oscillating or if there is no improvement and it is stuck around a suboptimal local optima, the learning rate is too high and should be decreased. On the contrary, if the learning is very slow and the loss does not decrease or if the model is overfitting, it is a clear sign that it should be increased.
The learning rate can take a wide range of values. Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored" \cite{28}. There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).\\
Batch size is another important hyperparameter to tune. Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples" \cite{29}. Furthermore, "networks train faster because the weights update is done after each propagation" \cite{29}. Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be" \cite{29}. Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.\\
Finally, the number of epochs during which the network is trained has to be carefully chosen. In fact, from a certain point in the training, neural networks don't learn anymore useful features in the data and start overfitting. This point corresponds to the moment where, the validation loss does not decrease anymore and starts to increase continuously. It is exactly at this moment that the training should stop. To achieve this goal, it is possible either to directly choose the right number of epochs or to use the so-called "early stopping" method, which stops the training as soon as the validation loss does not decrease for $N$ epochs.

@@

Convolutional Neural Networks (CNN) are a specific type of deep neural networks. They are particular in that they contain layers which perform a mathematical operation named convolution on the input data. CNNs are mostly used in image and video analysis.

>>>>

\chapter{Deep learning}
\label{ch:deep_learning}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work. It starts with a historical timeline of deep learning before describing what a neural network is. Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained. This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks". They are used in many computer vision applications due to their great performance on these tasks. Finally, the concept of "transfer learning" is discussed as the entire chapter \ref{ch:transfer_learning} relies on it.

\section{Introduction to deep learning}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Deep learning is currently one of the trendiest topics in machine learning, a subset of artificial intelligence. Machine learning refers to statistical models that allow computers to perform specific tasks without having been explicitly programmed to solve them. In fact, these models try to find structural patterns within data in order to understand new incoming situations and react in the best possible way. There exist various techniques in machine learning such as k-NN, SVM, k-means, decision trees, association rules, etc. What mainly differentiates deep learning from these algorithms is the concept of neural networks (see Section  \ref{what_is_a_neural_network}) that are combined to form deep neural networks.\\
Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, smarter translators, computer-aided diagnoses, personal assistants, art creation, robotics etc. The presence of deep learning techniques in these use cases clearly testifies to the enthusiasm of many areas for this technology. Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

\subsection{Historical background}
\label{historical_background}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}As described on figure \ref{history}, the theoretical foundations of deep learning appeared long before the invention of computers. From the first attempts to understand the human brain until today, huge progress was made to establish the basic components of modern neural networks. One could ask why deep learning took off recently if the theory was around for a long time.\\
As stated by Goodfellow et al. \cite{15}, the first part of the answer is computing power. In fact, deep learning algorithms need a lot of data to work properly, which requires powerful CPUs/GPUs that either didn't exist or were only within few people's reach. One other main reason concerns the lack of data. Since deep learning algorithms "learn" from data, learning is impossible if good-quality data is not available. The era of Big Data enhanced deep learning possibilities. 
%%%% Je mettrais soit l'un soit l'autre. Le paragraphe du dessus c'est le paragraphe du dessous avec des autres mots :)
% These two points are summarised: "The increase in model size overtime, due to the availability of faster CPUs, the advent of general purpose GPUs, fasternetwork, connectivity and better software infrastructure for distributed computing, is one of the most important trends in the history of deep learning. This trend is generally expected to continue well into the future".\\
Finally, before the year 2012, the abilities of neural networks were still to be proven. This changed with the ImageNet Large Scale Visual Recognition Challenge (a competition where researchers evaluate their algorithms on several visual recognition tasks). In fact, the deep convolutional neural network called "AlexNet" achieved 16\% of classification error rate, whereas the previous best scores were around 25\%. This victory marked the beginning of a new craze for these types of algorithms.

@@

\subsection{What is a neural network?}
\label{what_is_a_neural_network}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}From a descriptive point of view, neural networks can simply be seen as non-linear applications that associate an input to an output with respect to certain parameters. The input can be an image, a sound or any input that can be converted into numerical features. The output of a neural network depends on the problem it tries to solve. In computer vision, the most common types of outputs are classes (for classification problems) and pixel coordinates (for segmentation problems).\\
From a mathematical standpoint, a neural network can be defined as a non-linear function $f$ that associates an input $x$ to an output $y$ with respect to parameters $\theta$.
\begin{equation}
y = f(x,\theta)
\end{equation}
The parameters $\theta$ are estimated from the training samples.
%\begin{figure}[!h]
%\centering
%\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/neural_network.png}
%\caption{Neural Network example}
%\label{neural_network}
%\end{figure}

\subsection{Supervised and unsupervised learning}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Machine learning algorithms belongs can be classified in two classes. The first one is "supervised learning". It includes learning algorithms whose training samples are associated with their labels in order to find the optimal mapping between the input and the output. The second one is "unsupervised learning". In contrast to supervised algorithms, the latter rely on unlabeled data. Its main goal is to infer the natural structure  present in the data. Since the models presented in this work belong to the "supervised learning" category, notions explained below refer to this kind of algorithms.

@@

\subsection{Notation}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In order to keep the mathematical description of neural network consistent, this work will use Andrew Y. Ng's notation \cite{16}, who was a pioneer in deep learning.\\

@@

\subsection{Perceptrons}
\label{perceptron}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Perceptrons are the main components of neural networks. They were "developed in the 1950s and 1960s by the scientist Frank Rosenblatt, inspired by earlier work by Warren McCulloch and Walter Pitts" \cite{13}. Today, they are called "artificial neurons" or simply "neurons".\\
The output $a$ of a perceptron $j$ is a function $f$ of input $x=(x_{1}, ..., x_{n})$ weighted by a vector of weights $w_{}=(w_{1}, ..., w_{n})$, completed by a bias $b_{j}$ and associated to a non-linear activation function $g$:
\begin{equation}
\label{perceptron_equation}
a_{j} = f_{j}(x) = g((\sum_{k=1}^{n} x_{k} * w_{k}) + b_{j})
\end{equation}

@@

\subsection{Activation functions}
\label{activation_functions}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}
Once the computation of the weighted sum of all inputs for a specific neuron is done, the latter has to pass the sum through an activation function that will decide the proportion of the result that will be sent to the next layer. Activation functions must be non-linear in order to approximate extremely complex functions. In fact, neural networks are considered as universal approximators. Hornik et al. claim that "multilayer feedforward networks are capable of approximating any measurable function to any desired degree of accuracy, in a very specific and satisfying sense" \cite{17}. According to Thomas Epelbaum \cite{18}, the most commonly used activation functions are:\\\\
 \noindent \textbf{Sigmoid function}\\
 The sigmoid function is defined as:
 \begin{equation}
  g(x) = \frac{1}{1+e^{-x}}
 \end{equation}
 Its derivative is:
 \begin{equation}
 g'(x) = g(x)(1-g(x))
 \end{equation}

@@

\subsection{Multilayer perceptrons}
\label{multilayer_perceptron}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}A multilayer perceptron is a type of artifical neural networks. Du et al. \cite{23} define multilayer perceptrons as "feedforward networks with one or more layers of units between the
input and output layers" where "the output units represent a hyperplane in the space
of the input patterns."\\
A multilayer perceptron is composed of $L$ layers, each of them composed of $n^{l}_{h}$ perceptrons. The layers are organized in the following way:
\begin{itemize}
\item The input layer: It is the neural network entry point for the data. Generally speaking, the data is provided in the form of a matrix $X \in \R$ of size $(n_{x} \times batch\_size)$ with their corresponding labels $Y \in \R$ of size $(n_{y} \times batch\_size)$. The batch size defines the number of samples that will be given at the same time to the network and $n_{x}$ is the dimension of each sample. Moreover, $x^{(i)}$ is the $i^{th}$ sample represented as a column vector. The total number of training samples is given by $m$. Finally, $y^{(i)}$ is the output label for the $i^{th}$ example. For instance, suppose the number of samples is 100 and the batch size 32. In this situation, the network will be fed with 4 batches of sizes [32, 32, 32, 4] respectively.

@@

\section{Training a neural network}
\label{training_a_neural_network}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Training a neural network can be broken down into multiple steps. The first one is the "forward propagation" step. It consists in giving examples that need to be classified (or segmented, depending on the task) to the untrained neural network and to spread intermediate results through all layers of the network. %Passing all batches as input to a network is called an "epoch". 
After seeing every single batch, the loss is computed using a "loss function". The latter is used to evaluate the predictions of the neural network in comparison to their ground-truth. Then, the weights and biases of the networks are updated during a process called "backpropagation" in order find the global minimum of the loss function. As soon as the neural network has seen every single batch, the end of an "epoch" is reached. This process is repeated for a defined number of epochs.

\subsection{Forward propagation}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}The forward propagation is used to transmit the input through the entire neural network. Mathematically speaking, the forward propagation step for a specific layer $l$ is represented by two equations. The first equation denotes the weighted sum of the input given to the $l^{th}$ layer before passing through the activation function $g$: 
\begin{equation}
z^{[l]} = W_{x}^{[l]}x^{(i)} + b^{[l]}
\end{equation}
The second equation describes the effect of the activation function:
\begin{equation}
a^{[l]} = g^{[l]}(z^{[l]})
\end{equation}
Since the output of the activation function is then given as input to all the neurons of the next layer, the whole forward propagation step can be defined as:
\begin{equation}
a_{j}^{[l]} = g^{[l]} (\sum_{k} w_{jk}^{[l]}a_{k}^{[l-1]} + b_{j}^{[l]}) = g^{[l]} (z_{j}^{[l]}) 
\end{equation}
Figure \ref{forward_propagation} illustrates the computation of the forward propagation pass for the $l^{th}$ layer. The weight natrix $W_{jk}^{l}$ represents the weights associated with the connection between perceptron $k$ in layer $l$ and perceptron $j$ in layer $l+1$. This matrix is multiplied by the output of the previous layer $a_{j}^{[l-1]}$ before adding the bias $b^{[l]}$. The result is given as input to the activation function.

@@

\subsection{Loss computation}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}As stated by Thomas Epelbaum, "the loss function evaluates the error performed by the neural network when it tries to estimate the data to be predicted" \cite{18}. It is therefore useful to measure the penalty for a single input. On the contrary, when the goal is to get a more general overview of the error on the entire batch or on the entire dataset, the cost function $J$ is used. The latter is represented by $J(\hat{y}, y)$ where $\hat{y}$ is the prediction of the neural network and $y$ the real label.

@@

\subsection{Backpropagation}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Backpropagation relies on a technique called "gradient descent" to minimize the cost function $J(W, b)$. Generally speaking, "the intuition behind the backpropagation algorithm is as follows. Given a training example $(x^{(i)}, y^{(i)})$, we will first run a forward pass to compute all the activations throughout the network, including the output value of the network. Then, for each node $i$ in layer $l$, we would like to compute an "error term" $\partial^{(l)}_{i}$ that measures how much that node was "responsible" for any errors in our output. For an output node, we can directly measure the difference between the network’s activation and the true target value, and use that to define $\partial^{(n_{l})}_{i}$(where layer $n_{l}$ is the output layer). How about hidden units? For those, we will compute $\delta^{(l)}_{i}$ based on a weighted average of the error terms of the nodes that uses $a^{(l)}_{i}$ as an input" \cite{24}.

@@

\subsection{Metrics}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In classification tasks, four separate output labels can occur:
\begin{itemize}
\item True Positive (TP):  an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{correct}.
\item True Negative (TN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{correct}.
\item False Positive (FP): an output belongs to this class if the prediction that the latter \textbf{contains} a certain feature is \textbf{incorrect}.
\item False Negative (FN): an output belongs to this class if the prediction that the latter does \textbf{not contain} a certain feature is \textbf{incorrect}.
\end{itemize}
From these four categories, multiple metrics with their own specificities can be computed \cite{25}:
\begin{itemize}
\item Accuracy: "Ratio of the correctly labeled subjects to the whole pool of subjects".
\begin{equation}
Accuracy = \frac{(TP+TN)}{TP+FP+FN+TN}
\end{equation}
Accuracy is a great measure in the case of symmetric data (i.e the number of FN $\approx$ FP and their cost is similar). When this condition is not fulfilled, accuracy can lead to bad models. For instance, let's define a binary classification model that always outputs class 0. If the data is composed of 99 samples from class 0 and 1 sample from class 1, the accuracy is equal to 99\% but the model is not smart. Consequently, this metric has to be used in addition to other metrics.

@@

\subsection{Data}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}In deep learning, data is essential. As seen previously, neural networks learn features from it. Therefore, data has to be handled carefully and in the right way. Usually, it is split into three different sets:
\begin{itemize}
\item The training set: This is the part of the dataset that is used to train the neural network (the weights and biases).
\item The validation set: This dataset is used to evaluate a trained model. Usually, the evaluation on the validation set is performed every $N$ epochs, where $N$ is a fixed number. The validation set needs to come from the same distribution as the training set but should contain exclusively unseen data. This last point is crucial since the validation set shows how well the neural network generalizes on unknown data. The validation set can also be used as indicator to decide when the training should be stopped in order to prevent "overfitting" which is the behaviour of a model that fits to the training set too closely and don't generalize well. In fact, if the validation loss continuously increases for a certain number of epochs, going on with the training will increase the overfit. This fact of interrupting the training earlier is called "early stopping".
\item The test set: This last part of the dataset is used to establish the final evaluation of the model. It also contains unseen data exclusively. 
\end{itemize}
Regarding the way these three sets are split, it mostly depends on the number of samples available. If the latter is big, the data is split into training and testing using the ratio 80/20. Then the remaining training samples are also split into training and validation using the ratio 80/20. On the contrary, if there are few data available, k-fold cross-validation is a good practice. This technique consists in splitting the entire dataset into k folds. One fold is picked as test set and the others are considered as training sets. The model is trained on training folds and tested on the test set. Then, another test set is picked and the same process is repeated until all possible tests set are picked. At the end of the process, the results of all test sets are averaged, which provides a good estimation of the model's performance. This technique is summarized on figure \ref{cross_validation}.

@@

\subsection{Weight initialization}
\label{weight_initialization}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Before training a neural network, the weights have to be initialized in order to proceed to the first forward propagation. The initialization of the neural network weights is crucial since it will determine how quickly the network converges to an optimum. The idea behind weight initialization is to generate an initialization that "prevents layer activation outputs from exploding or vanishing during the course of a forward pass through a deep neural network. If either occurs, loss gradients will either be too large or too small to flow backwards beneficially, and the network will take longer to converge, if it is even able to do so at all" \cite{27}.\\
The simplest and least efficient technique to initialize neural network weights is to randomly generate them. The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions. Consequently, the slope of the gradient changes slowly and the learning takes a lot of time.\\
To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
\begin{equation}
\frac{\sqrt{6}}{\sqrt{n_{h}^{[l]}+n_{h}^{[l+1]}}}
\end{equation}
where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
For activation functions that are not symmetric around zero and don't have outputs inside [-1,1] such as ReLU or ELU, Kaiming initialization is an alternative. It consists in multiplying the randomly initialized  weight matrix by:
\begin{equation}
\frac{\sqrt{2}}{\sqrt{n_{h}^{[l]}}}
\end{equation} 
where $n_{h}^{[l]}$ is the number of incoming connections coming into a given layer from the previous layer's output.

\subsection{Hyperparameters tuning}

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Hyperparameters denote parameters that cannot be directly learned from the data. That is the case for the learning rate, the batch size and the number of epochs that were described in previous sections. So, these parameters have to be manually tuned in order to find the best configuration (i.e the one that minimizes the cost function and that keeps an acceptable level of generalization).\\
Regarding the learning rate $\alpha$, its value has to be neither too large nor too small. A too large learning rate is recognizable by analyzing the training loss curve: if the loss is exploding or oscillating or if there is no improvement and it is stuck around a suboptimal local optima, the learning rate is too high and should be decreased. On the contrary, if the learning is very slow and the loss does not decrease or if the model is overfitting, it is a clear sign that it should be increased.
The learning rate can take a wide range of values. Consequently, the most used technique to find the optimal learning rate is simply the "trial and error", which consists in "trying widely different learning rates to determine the range of learning rates that need to be explored" \cite{28}. There also exists methods that, instead of keeping a fixed learning rate for the entire training, reduce it after each epoch (learning rate decay) or each time the loss on the validation set does not decrease (learning rate scheduling).\\
Batch size is another important hyperparameter to tune. Training a network with a small batch size "requires less memory, since the latter is trained using fewer samples" \cite{29}. Furthermore, "networks train faster because the weights update is done after each propagation" \cite{29}. Nevertheless, "the smaller the batch the less accurate the estimate of the gradient will be" \cite{29}. Indeed, due to the high weights update frequency, the gradient fluctuates much more than if was computed after a bigger number of samples.\\
Finally, the number of epochs during which the network is trained has to be carefully chosen. In fact, from a certain point in the training, neural networks don't learn anymore useful features in the data and start overfitting. This point corresponds to the moment where, the validation loss does not decrease anymore and starts to increase continuously. It is exactly at this moment that the training should stop. To achieve this goal, it is possible either to directly choose the right number of epochs or to use the so-called "early stopping" method, which stops the training as soon as the validation loss does not decrease for $N$ epochs.

@@

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}Convolutional Neural Networks (CNN) are a specific type of deep neural networks. They are particular in that they contain layers which perform a mathematical operation named convolution on the input data. CNNs are mostly used in image and video analysis.
